% arara: pdflatex: { synctex: yes }
% arara: makeindex: { style: ctuthesis }
% arara: bibtex

% The class takes all the key=value arguments that \ctusetup does,
% and a couple more: draft and oneside
\documentclass[oneside]{ctuthesis}

%— PDF compression settings (lossless) —%
\pdfminorversion=5
\pdfcompresslevel=9
\pdfobjcompresslevel=3

\ctusetup{
	preprint = \ctuverlog,
	mainlanguage = english,
%	titlelanguage = czech,
	% mainlanguage = czech,
	otherlanguages = {slovak,english},
	title-czech = {Detailní vyhodnocení přístupů k úloze Structure-from-Motion},
	title-english = {A Detailed Evaluation of State-of-the-Art Structure-from-Motion Approaches},
	% subtitle-czech = {Cesta do tajů kdovíčeho},
	% subtitle-english = {Journey to the who-knows-what wondeland},
	doctype = M,
	faculty = F3,
	% department-czech = {Katedra matematiky},
	department-english = {Department of Cybernetics},
	author = {Valentin Lhermitte},
	supervisor = {Torsten Sattler, Dr. rer. nat.},
	% supervisor-address = {Ústav X, \\ Uliční 5, \\ Praha 99},
	% supervisor-specialist = {John Doe},
	fieldofstudy-english = {Open Informatics},
	subfieldofstudy-english = {Computer Vision and Image Proccessing},
	% fieldofstudy-czech = {Matematcké inženýrství},
	% subfieldofstudy-czech = {Matematické modelování},
	keywords-czech = {Structure-from-motion, Moderní SFM, Vyhodnocení polohy kamery, Syntéza nového pohledu, Vyhodnocovací protokol},
	keywords-english = {Structure-from-motion, Modern SFM, Camera pose evaluation, Novel view synthesis, Evaluation protocol},
	day = 23,
	month = 5,
	year = 2025,
	specification-file = {ctutest-zadani.pdf},
%	front-specification = true,
%	front-list-of-figures = false,
%	front-list-of-tables = false,
%	monochrome = true,
%	layout-short = true,
}

\ctuprocess

\addto\ctucaptionsczech{%
	\def\supervisorname{Vedoucí}%
	\def\subfieldofstudyname{Studijní program}%
}

\ctutemplateset{maketitle twocolumn default}{
	\begin{twocolumnfrontmatterpage}
		\ctutemplate{twocolumn.thanks}
		\ctutemplate{twocolumn.declaration}
		\ctutemplate{twocolumn.abstract.in.titlelanguage}
		\ctutemplate{twocolumn.abstract.in.secondlanguage}
		\ctutemplate{twocolumn.tableofcontents}
		\ctutemplate{twocolumn.listoffigures}
	\end{twocolumnfrontmatterpage}
}

% Theorem declarations, this is the reasonable default, anybody can do what they wish.
% If you prefer theorems in italics rather than slanted, use \theoremstyle{plainit}
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[chapter]
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{conjecture}[theorem]{Conjecture}

\theoremstyle{note}
\newtheorem*{remark*}{Remark}
\newtheorem{remark}[theorem]{Remark}

\setlength{\parskip}{2ex plus 0.2ex minus 0.2ex}

% Abstract in Czech
\begin{abstract-czech}
	Techniky SfM (Structure-from-Motion) jsou nedílnou součástí rekonstrukce 3D geometrie z 2D obrazových sekvencí.
	Nedávný pokrok v oblasti SfM vedl k vývoji několika nejmodernějších algoritmů, které využívají hluboké učení a další moderní techniky.
	Tato zpráva hodnotí čtyři současné algoritmy SfM \textbf{GLOMAP, VGGSfM, FlowMap} a \textbf{AceZero} na rozmanitých referenčních souborech dat.
	Hodnocení se zaměřuje na kritické ukazatele, včetně přesnosti odhadu polohy kamery, kvality syntézy nových pohledů a také časové efektivity a spotřeby paměti.
	Poznatky z této studie mají za cíl nasměrovat budoucí vývoj a použití SfM v počítačovém vidění a příbuzných oborech.
\end{abstract-czech}

% Abstract in English
\begin{abstract-english}
	Structure-from-Motion (SfM) techniques are integral in reconstructing 3D geometry from 2D image sequences.
	Recent advancements in SfM have led to the development of several state-of-the-art algorithms that leverage deep learning and other modern techniques.
	This report evaluates four contemporary SfM algorithms \textbf{GLOMAP, VGGSfM, FlowMap} and \textbf{AceZero} on a diverse set of benchmark datasets.
	The evaluation focuses on critical metrics, including camera pose estimation accuracy, the quality of novel view synthesis as well as time efficiency and memory consumption.
	The insights from this study aim to guide future development and application of SfM in computer vision and related fields.
\end{abstract-english}

% Acknowledgements / Podekovani
\begin{thanks}
	I would like to thank Torsten Sattler, Dr. rer. nat., for his guidance and supervision throughout this work.

	I would also like to acknowledge the support of the OP VVV funded project \text{CZ.02.1.01/0.0/0.0/16\_019/0000765} \text{``Research Center for Informatics''}
	which provided the necessary resources and infrastructure for this research.
\end{thanks}

% Declaration / Prohlaseni
\begin{declaration}
	I, Valentin Lhermitte, hereby declare that this report is a true and original work completed by myself, 
	under the guidance and supervision of Torsten Sattler, Dr. rer. nat.
	The work presented in this report has not been submitted elsewhere for assessment
	and has not been previously used to obtain any academic qualification or degree. All
	references and sources used in this report have been duly acknowledged and cited in the bibliography.
	I affirm that the data, facts, and opinions presented in this report are true and accurate
	to the best of my knowledge and belief. I take full responsibility for the content of this
	report and any errors or omissions therein.

In Prague, \ctufield{day}.~\monthinlanguage{title}~\ctufield{year}
\end{declaration}

% Only for testing purposes
\listfiles
\usepackage[pagewise]{lineno}
\usepackage{lipsum,blindtext}
\usepackage{mathrsfs} % provides \mathscr used in the ridiculous examples
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{subcaption} % Required for subfigure captions
\usepackage{float} % Required for H float positioning
\usepackage{array} % Required for bold table lines
\usepackage[dvipsnames]{xcolor}
\usepackage{colortbl}
% \usepackage[backend=biber,style=numeric]{biblatex}
\usepackage{hyperref}
\usepackage{booktabs}     
\usepackage{multirow}
\usepackage{longtable}
\usepackage{siunitx}         % aligns numbers on the decimal point
\usepackage{adjustbox}       % quick resize or clip
\usepackage{rotating} 	  	 % for sidewaystable
\usepackage{siunitx}
\usepackage{changepage}

\sisetup{table-format=1.2, detect-all, round-mode=places}

\hypersetup{
	colorlinks=true,
	linkcolor=red,
	citecolor=blue,
	urlcolor=blue,
}

\newcommand{\best}[1]{\textbf{#1}}
\newcommand{\sbest}[1]{\underline{#1}}

% \addbibresource{src/references.bib}

\begin{document}

\maketitle

\input{src/introduction.tex}
\input{src/related_work.tex}
\input{src/methodology.tex}
\input{src/experiments.tex}
\input{src/discussion.tex}
\input{src/conclusion.tex}

% \blindmathtrue

% \blinddocument

% \appendix

% \printindex

% \appendix

\bibliographystyle{plain}
\bibliography{src/references}
% \printbibliography

% \ctutemplate{specification.as.chapter}

\end{document}